---
title: "RWorksheet_Calambro#1"
author: "Christian Paul Calambro"
date: "2024-09-20"
output: pdf_document
---

1. Set up a Vector named age. 
```{r}
age<-c(34, 28, 22, 36,  27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 27, 34, 19, 20, 57, 49,50,37,46,25,17,37,42,53,41,51,35,24,33,41)
numofdatapts<-length(age)
numofdatapts
```
2. find the  reciprocal of the values of age.

```{r}
recip_age<-1/age
recip_age
```
3. Assign also new_age<-c(age, 0, age).
```{r}
new_age<-c(age, 0, age)
new_age
```
4. Sort the values for age.
```{r}
sort(age)
```
5. Find the minimun and maximun value  for age.
```{r}
min(age)
max(age)
```
6. Set up a Vector  named data.
```{r}
data<-c(2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5,  2.3, 2.4, 2.7)
numofdatapts<-length(data)
numofdatapts
```
7. Generate a new Vector for Data where you double every value  of the data.
```{r}
newdata<-2*data
newdata
```
8. Generate a sequence or the following scenario.
#*.1 Integers from 1-100.
```{r}
int <-seq(1:100)
int

#8.2 Numbers fron 20-60
num<-seq(from=20, to=60.)

#8.3 Mean of numbers from 20-60
num<-seq(from=20, to=60)
xnum<-mean(num)
xnum

#8.4 Sum of numbers from 51-91
num<-sum(51:91)
num

#8.5 Integers from 1-1000.
int<-seq(1:1000)
int

#a. How many data point from 8.1-8.4
datapts<-length(8.1:8.4)
datapts


#c. For 8.5, find only the maximum  data points until 10.
mxdatapts<-int[1:10]
mxdatapts
length(mxdatapts)
```

9. Print a vector with the integers between 1 and 100 that are not divisible by 3, 5, and 7 using filter option.
```{r}
Filter(function(i){all(i %%c(3,5,7) !=0)}, seq(100))
```

10. Generate a sequence backwards of the integers from 1-100.
```{r}
rev(seq(1:100))
```

11. List all the natural numbers below 25 that are multiples of 3 or 5.
```{r}
x<-Filter(function(i){any (i %% c(3,5) ==0)}, seq(24))
Sumofx<-sum(x)
Sumofx

#a. How many data points  from 10-11?
datapts<-length(10:11)
datapts
```

12. 
```{r}
#x<-{0+x+5+}(The code does not run because of a syntax error)
```

13. 
```{r}
score<-c(72, 86, 92, 63, 88, 89, 91, 92, 75, 75, 77)
s2<-score[2]
s2
s3<-score[3]
s3

```

14. Create  a vector a=c(1,2,NA,4,NA,6,7).
```{r}
a<-c(1, 2, NA, 4, NA, 6, 7)
print(a,na.print="-999")
```

15.
```{r}
name=readline(prompt="Input Your Name:")
age=readline(prompt="Input Your Age:")
print(paste("My name  is", name,"and I am", age, "years old"))
print(R.version.string)
```

