---
title: "RWorksheet_Calambro#4b"
author: "Christian Paul Calambro"
date: "2024-11-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#1
```{r}
vectorA <- c(1,2,3,4,5)
patmat <- matrix(0, nrow = 5, ncol = 5)

for (i in 1:5) {
  for (j in 1:5) {
    patmat[i, j] <- vectorA[(j + i - 2) %% 5 + 1]
  }
}
cat("Pattern Matrix:\n")
print(patmat)

zero_matrix <- matrix(0, nrow = 5, ncol = 5)
cat("\nZero Matrix:\n")
print(zero_matrix)
```
#2
```{r}
n <- 4


for (i in 1:n) {
  for (j in 1:i) {
    cat("*") 
  }
  cat("\n") 
}
```
#3
```{r}
first <- as.integer(readline(prompt = "Enter starting number for fibonacci sequence(up to 500): "))
fibonacci <- c(0,1)

repeat {
  next_fib <- sum(tail(fibonacci, 2))
  if (next_fib > 500) {
    break
  }
  fibonacci <- c(fibonacci, next_fib)
}

result <- fibonacci[fibonacci >= first]
cat("Fibonacci sequence from", first, "up to 500:", result, "\n")
```
#4a
```{r}
SHG <- read.csv("/cloud/project/Worksheet4/shoe_size_data.csv", header = TRUE, sep = ",")
SHG[1:6,]
```
#4b
```{r}
male_sub <- subset(SHG, Gender == "M")
female_sub <- subset(SHG, Gender == "F")

num_male <- nrow(male_sub)
num_female <- nrow(female_sub)

cat("Number of Male observations:", num_male, "\n")
cat("Number of Female observations:", num_female, "\n")
```
#4c
```{r}
gender_count <- table(SHG$Gender)
bar_colors <- c("blue", "pink")  # Colors for male and female
barplot(gender_count,
        main = "Number of Males and Females in Household Data",
        xlab = "Gender",
        ylab = "Number of Observations",
        col = bar_colors,
        legend = rownames(gender_count),
        beside = TRUE)

# Add grid lines
grid(nx = NULL, ny = NULL)
```
#5
```{r}

spending <- c(Food = 60, Electricity = 10, Savings = 5, Miscellaneous = 25)

percent_labels <- paste0(names(spending), ": ", round(spending / sum(spending) * 100, 1), "%")


colors <- c("lightblue", "lightgreen", "lightcoral", "lightpink")

pie(spending,
    labels = percent_labels,
    col = colors,
    main = "Dela Cruz Family Monthly Income Distribution"
)
```
#6a
```{r}
data(iris)
str(iris)
```
#6b
```{r}
mean_values <- colMeans(iris[, c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width")])
mean_values
```
#6c
```{r}
species_counts <- table(iris$Species)
pie(species_counts,
    labels = paste(names(species_counts), ": ", round(species_counts / sum(species_counts) * 100, 1), "%", sep=""),
    col = c("lightblue", "lightgreen", "lightcoral"),
    main = "Species Distribution in Iris Dataset"
)
```
#6d
```{r}
setosa_subset <- tail(subset(iris, Species == "setosa"), 6)
versicolor_subset <- tail(subset(iris, Species == "versicolor"), 6)
virginica_subset <- tail(subset(iris, Species == "virginica"), 6)

setosa_subset
versicolor_subset
virginica_subset

```
#6e
```{r}
plot(iris$Sepal.Length, iris$Sepal.Width,
     col = iris$Species,
     pch = 19,
     xlab = "Sepal Length",
     ylab = "Sepal Width",
     main = "Iris Dataset",
     sub = "Sepal width and length by Species"
)

legend("topright", legend = levels(iris$Species),
       col = 1:3, pch = 19)
```
#6f
#For part f, after making the scatterplot, you can explain what you see by looking at how the points are #grouped for each species. Notice if certain species have bigger or smaller sepals (the lengths and widths of #the flower parts). See if each type of flower is grouped together or spread out differently, which might #tell you how their sepal sizes vary from one species to another.

#7a
```{r}
library(readxl)  
alexa_data <- read_excel("/cloud/project/Worksheet4/alexa_file.xlsx")


alexa_data$variation <- gsub("Black\\s+Dot", "Black Dot", alexa_data$variation)
alexa_data$variation <- gsub("Black\\s+Plus", "Black Plus", alexa_data$variation)
alexa_data$variation <- gsub("Black\\s+Show", "Black Show", alexa_data$variation)
alexa_data$variation <- gsub("Black\\s+Spot", "Black Spot", alexa_data$variation)
alexa_data$variation <- gsub("White\\s+Dot", "White Dot", alexa_data$variation)
alexa_data$variation <- gsub("White\\s+Plus", "White Plus", alexa_data$variation)
alexa_data$variation <- gsub("White\\s+Show", "White Show", alexa_data$variation)
alexa_data$variation <- gsub("White\\s+Spot", "White Spot", alexa_data$variation)


table(alexa_data$variation)
```
#7b
```{r}
library(dplyr)


variations_count <- alexa_data %>%
  count(variation, name = "Total")


save(variations_count, file = "variations.RData")


print(variations_count)
```
#7c
```{r}

library(ggplot2)

load("variations.RData")


ggplot(variations_count, aes(x = variation, y = Total, fill = variation)) +
  geom_bar(stat = "identity") +
  ggtitle("Total Count of Alexa Variations") +
  xlab("Variation") +
  ylab("Total Numbers") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set3")
```
#7d
```{r}
variations_count$Category <- ifelse(grepl("Black", variations_count$variation), "Black Variants", 
                                    ifelse(grepl("White", variations_count$variation), "White Variants", NA))

black_white_variants <- variations_count %>% filter(!is.na(Category))


ggplot(black_white_variants, aes(x = variation, y = Total, fill = variation)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Category, scales = "free_x") +  
  ggtitle("Counts of Alexa Black and White Variants") +
  xlab("Variation") +
  ylab("Total Numbers") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set2")  
```