---
title: "RWorksheet_Calambro#4a"
author: "Christian Paul Calambro"
date: "2024-10-02"
output: pdf_document
---

1. The table below shows the data about shoe size and height. Create a data frame.
```{r}
library(readxl)
Household <- read_excel("HouseHold_Data.xlsx")

Household
```

a. Describe the data

The data set includes 25 people with information on their shoe size, height, and gender. Shoe sizes range from 5.0 to 13.0, and heights range from 59.0 to 77.0 inches. Males generally have larger shoe sizes and are taller than females. Taller individuals tend to have bigger shoe sizes.



b. Create a subset by males and females with their corresponding shoe size and height.
What its result? Show the R scripts.
```{r}
males <- subset(Household, Gender == "M", select = c(Shoe_size, Height))
females <- subset(Household, Gender == "F", select = c(Shoe_size, Height))
males
females
```




c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
result.
```{r}
meanofshoes <- mean(Household$Shoe_size)
meanofheight <- mean(Household$Height)

paste("The Mean of Shoes:", meanofshoes)
paste("The Mean of Height:", meanofheight)
```


d. Is there a relationship between shoe size and height? Why?
Yes, there is a relationship between the shoe size and the height.The taller individuals tend to have larger shoe sizes.



FACTORS

2. Construct character vector months to a factor with factor() and assign the result to
factor_months_vector. Print out factor_months_vector and assert that R prints out
the factor levels below the actual values.
```{r}
Months <- c("March", "April", "January", "November", "January",
            "September", "October", "September", "November", "August",
            "January", "November", "November", "February", "May", "August",
            "July", "December", "August", "August", "September", "November", 
            "February", "April")

Factor_Months <- factor(Months)
Factor_Months
levels(Months)
```


3. Then check the summary() of the months_vector and factor_months_vector. | Inter-
pret the results of both vectors. Are they both equally useful in this case?
```{r}
("The Summary of Months:")
summary(Months)


("The Summary of Factor_Months:")
summary(Factor_Months)
```



4. Create a vector and factor for the table below.
```{r}
Directions <- c("East", "West", "North")

Frequency <- c(1, 4, 3)

new_order_data <- factor(Directions, levels = c("East", "West", "North"))

("Directions Factor with Specified Order:")
new_order_data

```


5. Enter the data below in Excel with file name = import_march.csv
```{r}
library(readxl)
data <- read_excel("import_march.csv")

data
```


6. a. Create an R Program that allows the User to randomly select numbers from 1 to 50.
Then display the chosen number. If the number is beyond the range of the selected choice,
it will have to display a string “The number selected is beyond the range of 1 to 50”. If
number 20 is inputted by the User, it will have to display “TRUE”, otherwise display the
input number.

```{r}
Num <- readline(prompt = "Enter a number from 1-50:")
Num

```

If the number is beyond the range of the selected choice,
it will have to display a string “The number selected is beyond the range of 1 to 50”. If
number 20 is inputted by the User, it will have to display “TRUE”, otherwise display the
input number.

```{r}
Num <- readline(prompt = "Enter a number from 1-50:")
if(Num > 50){
  print("The number selected is beyond the range of 1 to 50.")
}else if(Num == 20){
  print("TRUE")
}else{print(Num)
}
```


7. Change
At ISATU University’s traditional cafeteria, snacks can only be purchased with bills. A
long-standing rule at the concession stand is that snacks must be purchased with as few
coins as possible. There are three types of bills: 50 pesos, 100 pesos, 200 pesos, 500 pesos,
1000 pesos.



a. Write a function that prints the minimum number of bills that must be paid, given the
price of the snack.
Input: Price of snack (a random number divisible by 50) Output: Minimum number of bills
needed to purchase a snack.

```{r}
min_amount <- function(price) {
  count <- 0
  
  if (price >= 1000) {
    count <- count + (price %/% 1000)
    price <- price %% 1000           
  }
  
  if (price >= 500) {
    count <- count + (price %/% 500)  
    price <- price %% 500          
  }
  
  if (price >= 200) {
    count <- count + (price %/% 200) 
    price <- price %% 200           
  }
  
  if (price >= 100) {
    count <- count + (price %/% 100)  
    price <- price %% 100             
  }
  
  if (price >= 50) {
    count <- count + (price %/% 50)  
    price <- price %% 50         
  }
  
  else(count)
}


price <- 1150  
cat("Minimum number of bills needed:", min_amount(price), "\n")

```



8. The following is each student’s math score for one semester. Based on this, answer the
following questions.



a. Create a dataframe from the above table. Write the R codes and its output.
```{r}
names <- c("Annie", "Thea", "Steve", "Hanna")
grade1 <- c(85, 65, 85, 100)
grade2 <- c(65, 75, 90, 90)
grade3 <- c(75, 55, 80, 85)
grade4 <- c(95, 75, 100, 90)

names_grades <- data.frame(Name = names , Grade1 = grade1, Grade2 = grade2, Grade3 = grade3, Grade4 = grade4)

print(names_grades)
```



b. Without using the rowMean function, output the average score of students whose average
math score over 90 points during the semester. write R code and its output.
Example Output: Annie’s average grade this semester is 88.75.
```{r}
for (i in 1:nrow(names_grades)) {

  average <- (names_grades$Grade1[i] + names_grades$Grade2[i] + names_grades$Grade3[i] + names_grades$Grade4[i]) / 4
  
  # Check if the average is over 90
  if (average > 90) {
    cat(names_grades$Name[i], "'s average grade this semester is ", average, ".\n", sep = "")
  }
}
```



c. Without using the mean function, output as follows for the tests in which the average
score was less than 80 out of 4 tests.
Example output: The nth test was difficult.
```{r}
for (test in 1:4) {

  total_score <- sum(names_grades[[paste0("Grade", test)]])
  

  average <- total_score / nrow(names_grades)
  
  
  if (average < 80) {
    cat("The", test, "test was difficult.\n")
  }
}
```



d. Without using the max function, output as follows for students whose highest score for
a semester exceeds 90 points.
Example Output: Annie’s highest grade this semester is 95.

```{r}
for (i in 1:nrow(names_grades)) {

  highest <- names_grades$Grade1[i]
  if (names_grades$Grade2[i] > highest) {
    highest <- names_grades$Grade2[i]
  }
  if (names_grades$Grade3[i] > highest) {
    highest <- names_grades$Grade3[i]
  }
  if (names_grades$Grade4[i] > highest) {
    highest <- names_grades$Grade4[i]
  }
  

  if (highest > 90) {
    cat(names_grades$Name[i], "'s highest grade this semester is ", highest, ".\n", sep = "")
  }
}
```






